---
title: "20% Criterion versus other Clinical Metrics"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Clear the environment
rm(list = ls())

# Free up memory by forcing garbage collection
invisible(gc())  

# Manually set the seed to an arbitrary number for consistency in reports
myseed <- 9

##_ Set knitr root directory to correspond to project working directory 
##_  setting based on structure with code in <project dir>/code
knitr::opts_knit$set(root.dir = here::here())
```

## Procedure

1. Load Patient Annotation
1. Scatter plot: MPCT against PFS for all subjects
1. Scatter plot: MPCT against PFS for baseline Affymetrix subjects
1. Waterfall: MPCT against 20pct classes,PDL1, Arm, BOR3
1. KM:  20pct classes versus PFS, OS

## Paths and Packages

```{r paths_packages}
# Provide paths
data_dir <- "./data/import"
results_dir <- "./results"
work_dir <- "./work"



## Load packages ##

#tidyverse  bundles: ggplot2, dplyr, tidyr, readr, purrr and tibble
suppressPackageStartupMessages(library(tidyverse))
library(ComplexHeatmap)
library(circlize)
library(ggpubr)
library(survival)
suppressPackageStartupMessages(library(survminer))
library(RColorBrewer)

```


```{r functions}

source("./code/ggplot_theme_dj_prm.R")
#' Custom ggplot theme with bolded text for easier legibility


```

```{r palettes}


# shape map for prior therapy
shape_prior <- c("Pretreated" = 16, 
              "Naive" = 18)


# color map for response
color_response <- c("Not20pct" = "black", 
              "20pctDec" = "goldenrod2")

```

## Data sources

```{r loaddata}

# Patient Annotation
# Anonymized
clinicalData_file <- paste(data_dir, "CM9_Patient_Annotation.txt", sep = "/" )
clinicalData <- read_tsv(clinicalData_file)

```

## Scatter plot for all Subjects

```{r scatter_all}

#make censoring a factor not numeric for plot purpose
#clinicalData$PFS_SOR <- factor(clinicalData$PFS_SOR, levels = c("0","1"))


plotcountscatter <- nrow(clinicalData %>% filter(Response20pct != "NE",
												 !is.na(MPCT)))

scatterplot_all <- clinicalData %>% filter(Response20pct != "NE",
										   !is.na(MPCT)) %>%
	ggplot() +
	geom_point(aes(x = PFS, y = MPCT, 
				   colour = Response20pct, shape = VEGFstatus),
			   size = 3, position = position_jitter(15)) +
	scale_colour_manual(name = 'Response',
						values = color_response) +
	scale_shape_manual(name = 'Prior Therapy',
						values = shape_prior) +
  geom_hline(yintercept = -20, linetype="dashed") +
  geom_text(data=data.frame(x=850,y=-25), 
  		  aes(x, y), label= ">20% Tumor Burden Reduction", size = 3) +
  #coord_fixed(ratio = 1) +
  labs(title = "CA209-009:\nMPCT versus PFS",
  	 subtitle = paste("Patients classified, N=",plotcountscatter),
       x = "PFS, days",
      y = "MPCT") +
  theme_dj(14) +
  theme(legend.position = "bottom") +
	guides(color=guide_legend(nrow=2,byrow=TRUE), shape = guide_legend(nrow=2,byrow=TRUE))

print(scatterplot_all)
```

## Scatter plot for Subjects with baseline Affy data

```{r scatter_affy}

#make censoring a factor not numeric for plot purpose
#clinicalData$PFS_SOR <- factor(clinicalData$PFS_SOR, levels = c("0","1"))


plotcountscatterbaseline <- nrow(clinicalData %>% filter(Response20pct != "NE" 
												 & !is.na(Subject_Affy_Status) 
												 	& Subject_Affy_Status != "13Post"))

scatterplotbaseline <- clinicalData %>% filter(Response20pct != "NE" 
												 & !is.na(Subject_Affy_Status) 
												 	& Subject_Affy_Status != "13Post") %>%
ggplot() +
	geom_point(aes(x = PFS, y = MPCT, 
				   colour = Response20pct, shape = VEGFstatus),
			   size = 3, position = position_jitter(15)) +
	scale_colour_manual(name = 'Response',
						values = color_response) +
	scale_shape_manual(name = 'Prior Therapy',
						values = shape_prior) +
  geom_hline(yintercept = -20, linetype="dashed") +
  geom_text(data=data.frame(x=850,y=-25), 
  		  aes(x, y), label= ">20% Tumor Burden Reduction", size = 3) +
  #coord_fixed(ratio = 1) +
  labs(title = "CA209-009:\nMPCT versus PFS",
  	 subtitle = paste("Patients with Baseline Affymetrix, N=",plotcountscatterbaseline),
       x = "PFS, days",
      y = "MPCT") +
  theme_dj(14) +
  theme(legend.position = "bottom") +
	guides(color=guide_legend(nrow=2,byrow=TRUE), shape = guide_legend(nrow=2,byrow=TRUE))

print(scatterplotbaseline)
```

## Waterfall plots with clinical annotation

```{r waterfall_select}

# This will be a zero data matrix
zero_row_mat = matrix(nrow = 0, ncol = nrow(clinicalData))
colnames(zero_row_mat) = clinicalData$MPCTrank


# Want to plot MPCT Response20pct PDL1cat BOR3 SUBJID
# Later plot TMB, VHL, PBRM1 SETD2?

# Provide column order as  = waterfall plot
ordering = order(-clinicalData$MPCTrank)

#Count subjects in plot
plotcount <- nrow(clinicalData)

ha_top = HeatmapAnnotation(barplot1 = anno_barplot(clinicalData$MPCT, 
												   axis = TRUE, 
												   baseline = 0,
						   						gp = gpar(fill = 
						   								  	ifelse(clinicalData$Response20pct == "20pctDec",
						   								  		   "goldenrod2", "black"))),
						   barplot2 = anno_barplot(clinicalData$MPCTbiopsyS1, 
												   axis = TRUE, 
												   baseline = 0,
						   						gp = gpar(fill = 
						   								  	ifelse(clinicalData$MPCTbiopsyS1 <= -20,
						   								  		   "goldenrod2", "black"))),
						   annotation_height = unit(c(3,3), "cm"))




ha_bottom = HeatmapAnnotation(response = clinicalData$Response20pct,
							  bor = clinicalData$BOR,
							  matchlesion = clinicalData$MatchedLesionBiopsy,
							  biopsy = clinicalData$Lesion_Site_Simple.S1,
							  col = list(matchlesion = c("TRUE" = "goldenrod2",
							  		   				"FALSE" = "black"),
							  		   biopsy = c("LymphNode"  = "goldenrod2",
							  		   		   "Adrenal" = "black","SoftTissue" = "black","Lung" = "black",
							  		   		   "Kidney"="black","Liver" = "black","Pancreas" = "black"),
							  		   response = c("Not20pct" ="black",
							  		   			 "20pctDec"= "goldenrod2",
							  		   			 "NE" = "white"),
							  bor = c("PD" = "black",
							  		"SD" = "grey",
							  		"PR" = "goldenrod1",
							  		"CR" = "goldenrod2",
							  		"NE" = "white")),
							  na_col = "white",
							  annotation_height = unit(c(.5,.5,.5,.5), "cm"))

#draw(ha_top,1:91)
#draw(ha_bottom,1:91)


heatmap_object = Heatmap(zero_row_mat, 
						 column_order = ordering,
						 column_title = "CA209-009: 91 subjects",
						 show_column_names = FALSE,
						 width = unit(11, "cm"),
						 top_annotation = ha_top,
						 top_annotation_height = unit(6, "cm"), 
						 bottom_annotation = ha_bottom,
						 bottom_annotation_height = unit(2, "cm"))



# Produce the heatmap and save to a file

heatmap_file <- paste(results_dir, "/Clinical_Heatmap_Annotation_",
					   plotcount,"_Subjects_Select.pdf",
					   sep="")
pdf(file=heatmap_file)


draw(heatmap_object, padding = unit(c(4, 10, 4, 4), "mm"))

# Decorate the heatmap object, adding layers of texts or lines

decorate_annotation("barplot1", 
					{grid.text("Tumor Burden\n\ Reduction",
							   unit(-10, "mm"), just = "bottom",
							   rot = 90, check.overlap = T,
							   gp = gpar(fontsize = 8))})
decorate_annotation("barplot2", 
					{grid.text("Baseline Biopsy\n\ Reduction",
							   unit(-10, "mm"), just = "bottom",
							   rot = 90, check.overlap = T,
							   gp = gpar(fontsize = 8))})

decorate_annotation("matchlesion", 
					{grid.text("Lesion Pair",
							   unit(-2, "mm"), just = "right",
							   gp = gpar(fontsize = 10))})

decorate_annotation("biopsy", 
					{grid.text("Biopsy Site",
							   unit(-2, "mm"), just = "right",
							   gp = gpar(fontsize = 10))})


decorate_annotation("response", 
					{grid.text("Response",
							   unit(-2, "mm"), just = "right",
							   gp = gpar(fontsize = 10))})

decorate_annotation("bor", 
					{grid.text("RECIST BOR",
							   unit(-2, "mm"), just = "right",
							   gp = gpar(fontsize = 8))})




dev.off()

```

```{r waterfall_all}

# This will be a zero data matrix
zero_row_mat = matrix(nrow = 0, ncol = nrow(clinicalData))
colnames(zero_row_mat) = clinicalData$MPCTrank


# Want to plot MPCT Response20pct PDL1cat BOR3 SUBJID
# Later plot TMB, VHL, PBRM1 SETD2?

# Provide column order as  = waterfall plot
ordering = order(-clinicalData$MPCTrank)

#Count subjects in plot
plotcount <- nrow(clinicalData)

ha_top = HeatmapAnnotation(barplot1 = anno_barplot(clinicalData$MPCT, 
												   axis = TRUE, 
												   baseline = 0,
						   						gp = gpar(fill = 
						   								  	ifelse(clinicalData$Response20pct == "20pctDec",
						   								  		   "goldenrod2", "black"))),
						   barplot2 = anno_barplot(clinicalData$MPCTbiopsyS1, 
												   axis = TRUE, 
												   baseline = 0,
						   						gp = gpar(fill = 
						   								  	ifelse(clinicalData$MPCTbiopsyS1 <= -20,
						   								  		   "goldenrod2", "black"))),
						   barplot3 = anno_barplot(clinicalData$TMB,
						   					axis = TRUE, 
												   baseline = 0),
						   annotation_height = unit(c(3,3,2), "cm"))




ha_bottom = HeatmapAnnotation(response = clinicalData$Response20pct,
							  bor = clinicalData$BOR,
							  pdl1 = clinicalData$PDL1cat,
							  vegf = clinicalData$VEGFstatus,
							  matchlesion = clinicalData$MatchedLesionBiopsy,
							  biopsy = clinicalData$Lesion_Site_Simple.S1,
							  AffyStatus = clinicalData$Subject_Affy_Status,
							  colname = anno_text(clinicalData$SUBJID,
												  rot = -90, just = "left", offset = unit(1, "npc") - unit(2, "mm"),
												  gp = gpar(fontsize = 7)),
							  col = list(vegf = c("Pretreated"="goldenrod2",
							  					"Naive"="darkolivegreen4"),
							  		   pdl1 = c("0%" = "grey", 
							  					"1to5%" = "red",
							  					"5+%" = "red"),
							  		   matchlesion = c("TRUE" = "goldenrod2",
							  		   				"FALSE" = "black"),
							  		   biopsy = c("LymphNode"  = "goldenrod2",
							  		   		   "Adrenal" = "black","SoftTissue" = "black","Lung" = "black",
							  		   		   "Kidney"="black","Liver" = "black","Pancreas" = "black"),
							  		   AffyStatus = c("17Pre"="darkolivegreen4",
							  						  "13Post" = "blue",
							  						  "42Pairs"="goldenrod2"),
							  		   response = c("Not20pct" ="black",
							  		   			 "20pctDec"= "goldenrod2",
							  		   			 "NE" = "white"),
							  bor = c("PD" = "black",
							  		"SD" = "grey",
							  		"PR" = "goldenrod1",
							  		"CR" = "goldenrod2",
							  		"NE" = "white")),
							  na_col = "white",
							  annotation_height = unit(c(.5,.5,.5,.5,.5,.5,.5,.5), "cm"))

draw(ha_top,1:91)
draw(ha_bottom,1:91)


heatmap_object = Heatmap(zero_row_mat, 
						 column_order = ordering,
						 column_title = "CA209-009: 91 subjects",
						 show_column_names = FALSE,
						 width = unit(11, "cm"),
						 top_annotation = ha_top,
						 top_annotation_height = unit(9, "cm"), 
						 bottom_annotation = ha_bottom,
						 bottom_annotation_height = unit(4, "cm"))



# Produce the heatmap and save to a file

heatmap_file <- paste(results_dir, "/Clinical_Heatmap_Annotation_",
					   plotcount,"_Subjects_All.pdf",
					   sep="")
pdf(file=heatmap_file)


draw(heatmap_object, padding = unit(c(4, 10, 4, 4), "mm"))

# Decorate the heatmap object, adding layers of texts or lines

decorate_annotation("barplot1", 
					{grid.text("Tumor Burden\n\ Reduction",
							   unit(-10, "mm"), just = "bottom",
							   rot = 90, check.overlap = T,
							   gp = gpar(fontsize = 8))})
decorate_annotation("barplot2", 
					{grid.text("Baseline Biopsy\n\ Reduction",
							   unit(-10, "mm"), just = "bottom",
							   rot = 90, check.overlap = T,
							   gp = gpar(fontsize = 8))})


decorate_annotation("barplot3", 
					{grid.text("TMB",
							   unit(-10, "mm"), just = "bottom",
							   rot = 90, check.overlap = T,
							   gp = gpar(fontsize = 9))})

decorate_annotation("vegf", 
					{grid.text("Prior VEGFi",
							   unit(-2, "mm"), just = "right",
							   gp = gpar(fontsize = 10))})
decorate_annotation("matchlesion", 
					{grid.text("Lesion Pair",
							   unit(-2, "mm"), just = "right",
							   gp = gpar(fontsize = 10))})

decorate_annotation("biopsy", 
					{grid.text("Biopsy Site",
							   unit(-2, "mm"), just = "right",
							   gp = gpar(fontsize = 10))})

decorate_annotation("pdl1", 
					{grid.text("PDL1 category",
							   unit(-2, "mm"), just = "right",
							   gp = gpar(fontsize = 8))})

decorate_annotation("response", 
					{grid.text("Response",
							   unit(-2, "mm"), just = "right",
							   gp = gpar(fontsize = 10))})

decorate_annotation("bor", 
					{grid.text("BOR",
							   unit(-2, "mm"), just = "right",
							   gp = gpar(fontsize = 10))})

decorate_annotation("AffyStatus", 
					{grid.text("Affymetrix",
							   unit(-2, "mm"), just = "right",
							   gp = gpar(fontsize = 10))})


dev.off()

```
## PFS: Median by 20% group 

```{r median_pfs}

clinicalData2 = filter(clinicalData, Response20pct != "NE")

pfs_cont <- survfit(Surv(clinicalData2$PFS, 1-clinicalData2$PFS_SOR) ~ Response20pct,
					data = clinicalData2, conf.type=c("log-log"))

df <- as.data.frame(summary(pfs_cont)$table[, c("n.start", "events", "median", "0.95LCL", "0.95UCL" )])
rn <- gsub( "Response20pct=", "", rownames(df))

as.data.frame(df, row.names = rn)

```

## PFS: KM by 20% group

```{r KM_PFS}

#make a 2-level palette
two_palette <-c( RColorBrewer::brewer.pal(9,"Blues")[c(5,9)],
                 RColorBrewer::brewer.pal(9,"Reds")[c(5,9)],
                 RColorBrewer::brewer.pal(9,"Greens")[c(5,9)])
				 
#Count subjects in plot
plotcount <- nrow(clinicalData2)
                 
# Produce the survfit object created in chunk {r median_pfs} and save to a file

pfs_file <- paste(results_dir, "/Clinical_KM_PFS_",
					   plotcount,"_Subjects.pdf",
					   sep="")
pdf(file=pfs_file)

ggsurvplot(pfs_cont,
		   legend.title = "Response category:",
		   legend.labs = c("Responder", "NonResponder"),
		   risk.table = T, pval=T, conf.int=T, risk.table.y.text.col=T, risk.table.y.text=T,
		   palette = c("goldenrod2", "black"),
		   surv.median.line = c("h"),
		   title= "CA209-009: PFS by 20% Response Category",
		   subtitle = paste0("Categorized Subjects, N = ",plotcount))+ 
	guides(colour = guide_legend(nrow = 2))

dev.off()
```

## OS: Median by 20% group 

```{r median_os}

clinicalData2 = filter(clinicalData, Response20pct != "NE")

os_cont <- survfit(Surv(clinicalData2$OS, 1-clinicalData2$OS_SOR) ~ Response20pct + VEGFstatus,
					data = clinicalData2, conf.type=c("log-log"))

df <- as.data.frame(summary(os_cont)$table[, c("n.start", "events", "median", "0.95LCL", "0.95UCL" )])
rn <- gsub( "Response20pct=", "", rownames(df))
rn <- gsub( "VEGFstatus=", "", rn)

as.data.frame(df, row.names = rn)

```

## OS: KM by 20% group

```{r KM_OFS}

#make a 2-level palette
two_palette <-c( RColorBrewer::brewer.pal(9,"Oranges")[c(5,9)],
                 RColorBrewer::brewer.pal(9,"Greys")[c(5,9)],
                 RColorBrewer::brewer.pal(9,"Greens")[c(5,9)])


#Count subjects in plot
plotcount <- nrow(clinicalData2)
                 
# Produce the survfit object created in chunk {r median_pfs} and save to a file

os_file <- paste(results_dir, "/Clinical_KM_OS_",
					   plotcount,"_Subjects.pdf",
					   sep="")
pdf(file=os_file)

ggsurvplot(os_cont,
		   legend.title = "Response category:",
		   legend.labs = c("Responder, Naive", "Responder, Pretreated",
		   				"NonResponder, Naive", "NonResponder, Pretreated"),
		   risk.table = T, pval=T, conf.int=F, risk.table.y.text.col=T, risk.table.y.text=T,
		   palette = two_palette,
		   surv.median.line = c("h"),
		   title= "CA209-009: OS by 20% Response Category",
		   subtitle = paste0("Categorized Subjects, N = ",plotcount)) + 
	guides(colour = guide_legend(nrow = 4))

dev.off()
```

## Outputs

```{r outputs}

# Save scatter plot
scatter_file <- paste(results_dir, "/Clinical_Scatterplot_MPCT_PFS_",
					   plotcountscatter,"_Subjects.png",
					   sep="")

ggsave(scatterplot_all, file = scatter_file, width=7, height=5,
	   units = "in", dpi = 96)

# Save scatter plot
scatterbaseline_file <- paste(results_dir, "/Clinical_Scatterplot_MPCT_PFS_",
					   plotcountscatterbaseline,"_Subjects_Baseline.png",
					   sep="")

ggsave(scatterplotbaseline, file = scatterbaseline_file, width=7, height=5,
	   units = "in", dpi = 96)


```

