---
title: "Genes Change over time associated with Response"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Clear the environment
rm(list = ls())

# Free up memory by forcing garbage collection
invisible(gc())  

# Manually set the seed to an arbitrary number for consistency in reports
myseed <- 9

##_ Set knitr root directory to correspond to project working directory 
##_  setting based on structure with code in <project dir>/code
knitr::opts_knit$set(root.dir = here::here())
```

## Procedure

1. Load rma data, nlme results, sample annotation, probeset annotation
1. Load clinical data file, used to annotate 42 paired sample Patients
1. Filter nlme results to 189 genes associated with response change over time
1. AND with >1.25-fold change in responders
1. I am not using criteria for a difference at Week 4 (56prbs)
1. Calculate the effect size
1. Subset the RMA data to the 42 paired samples
1. Make dataframe of Diff in RMA values (Week4 - Screen) per Patient
1. Make Annotated Heatmap of Change in Expression
1. Make boxplot of top hit
1. Save table of 189 genes

# Paths and Packages

```{r paths_packages}
# Provide paths
data_dir <- "./data/import"
results_dir <- "./results"
work_dir <- "./work"


## Load packages ##
#tidyverse  bundles: ggplot2, dplyr, tidyr, readr, purrr and tibble
suppressPackageStartupMessages(library(tidyverse))
library(ComplexHeatmap)
library(circlize)
library(ggpubr)
```



```{r functions}

source("./code/ggplot_theme_dj_prm.R")
#' Custom ggplot theme with bolded text for easier legibility


```

```{r palettes}


# shape map for prior therapy
shape_prior <- c("PriorTherapy_Yes" = 16, 
              "PriorTherapy_No" = 18)


# color map for response
color_response <- c("Not20pct" = "black", 
              "20pctDec" = "goldenrod2")

```

## Data Sources

```{r loaddata}

# nlme Results
TumorRESPfile <- paste(data_dir, "CA209009-tumorAffy-table02-v01.csv", sep = "/")
BiopsyRESP <- read.csv(TumorRESPfile, stringsAsFactors=FALSE, header=TRUE, na.strings = "NA")

# SDRF (Sample and Data Relationship Format) file from Array Express
sdrf_file <- paste(data_dir, "E-MTAB-3218 sdrf_response.txt", sep = "/")
sdrf <- read_tsv(sdrf_file)

# Affymetrix probeset to Gene annotation
probeset_file <- paste(data_dir, "U219_BrainArray_Locus_Symbol_IRIS.txt", sep = "/")
probeset <- read_tsv(probeset_file)

# Expression values
rma_file <- paste(data_dir, "CA209009-tumorAffy-HGU219_HS_ENTREZG.rma", sep = "/")
rma <- read_tsv(rma_file)


# Patient Annotation
# Anonymized
clinicalData_file <- paste(data_dir, "CM9_Patient_Annotation.txt", sep = "/" )
clinicalData <- read_tsv(clinicalData_file)


```

## Time.response Gene Results Table

I am going to select the 189 genes using:

+ p.response < P05 (Umbrella test for difference in means at any timepoint)
+ p.response.time < P01 (Contrast for difference in slope of change over time)
+ absolute difference in Responder means  >= 0.32165 (1.25-fold effect size over Time in Responders)


```{r results}
# Filter Parameters
P01 <- 0.01
P05 <- 0.05
Fold13 <- 0.3785
Fold12 <- 0.263
Fold125 <- 0.32165


# To subset the 189 probesets with a Time.Response effect, with 1.25x in Responders
TumTimeRESPeffect <- filter(BiopsyRESP, 
							p.response < P05,
							p.response.time < P01,
							#p.response28 < P01,
							#abs(response28) > Fold125
							abs(DecGE20pc_Day28_GrpAvg-DecGE20pc_Day0_GrpAvg) > Fold125) %>% 
	dplyr::select(Gene, p.response, 
				  p.response.time, q.response.time,
				  p.response28,
		   DecLT20pc_Day0_GrpAvg, DecGE20pc_Day0_GrpAvg,
		   DecLT20pc_Day28_GrpAvg, DecGE20pc_Day28_GrpAvg)


# Merge in updated gene annotation
TumTimeRESPeffect <- right_join(probeset [,1:5], TumTimeRESPeffect,
								  by = c("Probeset" = "Gene"))

# calculate Effect size using modeled group means
TumTimeRESPeffect <- mutate(TumTimeRESPeffect, 
									 Effect.Size.Resp = round(DecGE20pc_Day28_GrpAvg-DecGE20pc_Day0_GrpAvg,2),
									Effect.Size.NonResp = round(DecLT20pc_Day28_GrpAvg-DecLT20pc_Day0_GrpAvg,2))


# calculate Foldchange using modeled group means
TumTimeRESPeffect <- mutate(TumTimeRESPeffect, 
								   Fold.Change.Resp = round(2^(DecGE20pc_Day28_GrpAvg-DecGE20pc_Day0_GrpAvg),2),
								   Fold.Change.NonResp = round(2^(DecLT20pc_Day28_GrpAvg-DecLT20pc_Day0_GrpAvg),2))


# Round modeled group mean RMA values
TumTimeRESPeffect <- TumTimeRESPeffect %>%
		mutate(Responder_Day0_GrpAvg = round(DecGE20pc_Day0_GrpAvg,2))%>%
	mutate(NonResponder_Day0_GrpAvg = round(DecLT20pc_Day0_GrpAvg,2))%>%
	mutate(Responder_Day28_GrpAvg = round(DecGE20pc_Day28_GrpAvg,2))%>%
	mutate(NonResponder_Day28_GrpAvg = round(DecLT20pc_Day28_GrpAvg,2))%>%
	select(-DecGE20pc_Day0_GrpAvg, -DecLT20pc_Day0_GrpAvg,
		   -DecGE20pc_Day28_GrpAvg, -DecLT20pc_Day28_GrpAvg)


```

## Time.Response Subject sample Annotation

Subset the Sample annotation to the 42 paired samples

```{r samples}

sdrfTime <- sdrf %>%
	filter(Subject_Affy_Status == "42Pairs", Response20pct != "NE")

annot42 <- clinicalData %>%
	filter(Subject_Affy_Status == "42Pairs", Response20pct != "NE")

# Insert MPCT biopsy where the lesion is the same pre/post
annot42$MPCTbiopsy <- annot42$MPCTbiopsyS1
annot42$MPCTbiopsy[annot42$MatchedLesionBiopsy == "FALSE"] <- NA

```

## Screen and Week4 Gene RMA expression Values

Obtain the RMA data for only the paired samples

```{r rma_diff}

# Make Colnames match "Assay.Name" in sdrf annotation
colnames(rma) <- sub("ENTREZG-", "", colnames(rma))
rma <- dplyr::rename(rma, Probeset = X1)

# Filter rma data to 189 probesets and 84 paired Array columns plus Probeset column
rma189 <- rma[rma$Probeset %in% TumTimeRESPeffect$Probeset, ]
rma189 <- select(rma189, one_of("Probeset",sdrfTime$Assay.Name))

# Each observation in one row
# 189 probesets x 84 observations
rma189rows <- rma189 %>%
	gather(key = "Assay.Name", value = "RMA", -Probeset) %>%
	left_join(select(sdrf, 
					 c("Assay.Name", "biopsy.timepoint", "Response20pct", 
					   "individual", "clinical.history"))) 

# Make dataframe of diff in RNA values (Week4 - Screen) per subject
# 189 probesets by 42 subjects
rma189diff <- rma189rows %>%
	select(-Assay.Name) %>% #if not there are 2 assays per subject per probeset
	group_by(individual,Probeset) %>% #one row per subject per probeset
	spread(biopsy.timepoint,RMA) %>% #2 timepoint columns with RMA value
	mutate(Change = Week4 - Screen) %>% #caculate Diff
	select(Probeset,individual,Change) %>% #only necessary columns
	group_by(Probeset) %>% #one row per probeset
	spread(individual,Change) #42 Individuals columns with Change value

#Put data columns in the same order as the annotation dataframe
colorder <- annot42$SUBJID
rma189diff <- rma189diff[,c("Probeset",colorder)]
```

## Heatmap of the change in expression values

```{r heatmap}

#make a row annotation matrix in the rma datset order
row189 <- left_join(rma189diff[,1], probeset[,1:5], by = "Probeset")

# Matrix will be the heatmap of Diff values (log fold change)
mat = rma189diff[,-1]
rownames(mat) =  rma189diff$Probeset


# Top heatmap annotation is tumor burden reduction; MPCT and MPCT biopsy
ha_top = HeatmapAnnotation(barplot2 = anno_barplot(annot42$MPCTbiopsy, axis = TRUE,baseline = 0,
						   						gp = gpar(fill = ifelse(annot42$MPCTbiopsy > -19,
						   												"black", "goldenrod2"))),
						      annotation_height = unit(c(3), "cm"))

#draw(ha_top,1:42)

# Bottom heatmap annotation is response annotation, BOR?
ha_bottom = HeatmapAnnotation(response = annot42$Response20pct,
							  Matched = annot42$MatchedLesionBiopsy,
							  colname = anno_text(annot42$SUBJID, rot = 90, just = "right", offset = unit(1, "npc") - unit(2, "mm"), 
							  					gp = gpar(fontsize = 8)),
							  col = list(Matched = c("TRUE"="goldenrod2","FALSE"="Black"),
							  		   response = c ("Not20pct" ="black", "20pctDec"= "goldenrod2")),
							  na_col = "white",
							  annotation_height = unit(c(0.5, 0.5,1), "cm"))

# draw(ha_bottom,1:42)

# Provide column order as descending MPCT (ie waterfall plot)
ordering = order(-annot42$MPCTrank)

heatmap_object = Heatmap(mat,
						 col = colorRamp2(c(-1, 0, 1), c("dodgerblue", "white", "firebrick")), 
						 cluster_rows = TRUE,
						 km = 2,
						 show_row_names = FALSE,
						 cluster_columns = FALSE,
						 column_order = ordering,
						 column_title = "CA209-009: 42 pairs for 189 Response.Time",
						 show_column_names = FALSE,
						 width = unit(10, "cm"),
						 top_annotation = ha_top,
						 top_annotation_height = unit(3, "cm"), 
						 bottom_annotation = ha_bottom,
						 bottom_annotation_height = unit(2, "cm")) +
	rowAnnotation(IRIS = row189$IRIS_Most_Specific,
				  col = list(IRIS = 
				  		   	c(`T Cell` = "darkolivegreen1",
				  		   	  `NK Cell` = "darkolivegreen1",
				  		   	  `B Cell` = "darkolivegreen1",
				  		   	  Lymphoid = "darkolivegreen1",
				  		   	  `Dendritic Cell` = "blue",
				  		   	  Neutrophil = "blue",
				  		   	  Monocyte = "blue",
				  		   	  Myeloid = "blue",
				  		   	  Multiple = "goldenrod2")),
				  na_col = "white",
				  width = unit(0.5, "cm"),
				  show_annotation_name = TRUE)



heatmap_file <- paste0(results_dir,"/","GEP_Heatmap_Biopsy_All_TimeResponseEffect.pdf")

pdf(file=heatmap_file)



draw(heatmap_object, padding = unit(c(2, 20, 2, 2), "mm"))

# Decorate the heatmap object, adding layers of texts or lines
decorate_annotation("Matched", 
					{grid.text("Matched Lesion",
							   unit(-2, "mm"), just = "right",
							   gp = gpar(fontsize = 8))})
decorate_annotation("response", 
					{grid.text("Response",
							   unit(-2, "mm"), just = "right",
							   gp = gpar(fontsize = 10))})

# decorate_annotation("barplot1", 
# 					{grid.text("MPCT",
# 							   unit(-10, "mm"), just = "bottom",
# 							   rot = 90, check.overlap = T,
# 							   gp = gpar(fontsize = 10))})


decorate_annotation("barplot2", 
					{grid.text("Matched Biopsy\n\ Reduction",
							   unit(-10, "mm"), just = "bottom",
							   rot = 90, check.overlap = T,
							   gp = gpar(fontsize = 10))})

	
dev.off()

```

## Boxplot of a selected Probeset

```{r boxplot}

mycomparisons <- list(c("Screen","Week4"))

# Define the desired probeset and get its Symbol
#3606_at LOC3606 IL18
probesetwanted <- "3606_at"
genewanted <- as.character(probeset[probeset$Probeset == probesetwanted, 4])

# Get Gene values, then transpose to Array ID and RMA values in columns
onegene <- rma %>% 
	filter(rma$Probeset == probesetwanted) %>%
	gather(key = "Assay.Name", value = "RMA", -Probeset) %>%
	left_join(select(sdrf, 
					 c("Assay.Name", "biopsy.timepoint", "Response20pct", 
					   "individual", "clinical.history"))) %>%
	filter(Response20pct != "NE")

# Set levels of Response factor to determine plotting order
onegene$Response20pct <-factor(onegene$Response20pct, 
							   levels = c("Not20pct","20pctDec"))
# Set levels of biopsy.timepoint factor to determine plotting order
onegene$biopsy.timepoint <-factor(onegene$biopsy.timepoint, 
							   levels = c("Screen","Week4"))

# Convert "individual" from numeric
onegene$individual <- as.factor(as.character(onegene$individual))

# Count number of datapoints plotted
plotcount <- nrow(onegene)

# Plot baseline vs Week 4 expression values for gene
boxplot <- ggplot(onegene,
				  aes(x = biopsy.timepoint, y = RMA)) +
	geom_boxplot(aes(biopsy.timepoint),
				 outlier.shape = NA,
				 alpha=0.2) +
	geom_point(aes(color = Response20pct,shape= clinical.history),
			   size = 2,
			   position = position_jitter(w = 0.2, h = 0)) +
		scale_colour_manual(name = 'Response',
						values = c("black", "goldenrod2")) +
	 scale_shape_manual(name = 'Prior Therapy',
	 						values=c(16,18)) +
	geom_line(aes(group = individual), 
			  colour = "black", alpha=0.3) +
			 scale_y_continuous(breaks=seq(3,9,1),
		 			   limits=c(3, 9))+
	facet_grid(~Response20pct)+
	labs(title = paste("CA209-009 Biopsy: ",probesetwanted, "=", genewanted),
		 subtitle = paste("Screen and/or Week4 Affymetrix, N=",plotcount),
		 x = "Response Category",
		 y = "Expression signal, RMA") +
	theme(legend.position = "bottom",
		  axis.text=element_text(size=12),
		  axis.title=element_text(size=14,face="bold")) +
	guides(shape = guide_legend(nrow=2,byrow=TRUE),
		   fill = guide_legend(nrow=2,byrow=TRUE)) +
				scale_x_discrete(labels=c("Screen" = "Screen", 
							  "Week4" = "Day28"))+
	stat_compare_means(method="t.test", size = 6,
					   aes(label = paste0("P = ", ..p.format..)),
					   comparisons = mycomparisons,
					   vjust = "inward")+
	theme_dj(16)

print(boxplot)
```

##  Output Results

```{r output}


# Table of 189 genes
# Count the hits and make an output filename
countTimeRESP <- nrow(TumTimeRESPeffect)
TumTimeRESP_file <- paste0(results_dir,"/",
						"GEP_Table_BiopsyAll_TimeResponseEffect_",countTimeRESP,"prbs",".txt")

# Write tsv file of 189prbs, annotation and statistics 
write_tsv(TumTimeRESPeffect, TumTimeRESP_file)

# Save box plot of selected gene
box_file <- paste(results_dir, "/GEP_Boxplot_Time_By_Response_",
				  paste0(probesetwanted,"_",genewanted),
				  ".png",
				  sep="")

ggsave(boxplot, file = box_file, width=7, height=4,
	   units = "in", dpi = 96)

```



