---
title: "CA209009 - Gene Set Enrichment - *CD3TCRsig_Baseline_Day28*"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Clear the environment
rm(list = ls())

# Free up memory by forcing garbage collection
invisible(gc())  

# Manually set the seed to an arbitrary number for consistency in reports
myseed <- 8675309  # Jenny, I've got your number
```

```{r set_wd_interactive, include=FALSE}
if ("rstudioapi" %in% installed.packages()[, "Package"] & rstudioapi::isAvailable() & interactive()) {
  # When in RStudio, dynamically sets working directory to path of this script
  setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
}
```

## Procedure

1. Load precomputed `limma` fit objects
1. Compute gene set enrichments with the `limma::cameraPR` method using [MSigDB collections](http://software.broadinstitute.org/gsea/msigdb/collections.jsp) *C2 curated gene sets* and *Hallmark gene sets*.
1. After enrichment, filter to showcase only those gene sets that are net down-regulated
1. Gene set enrichment plots and tables of *net downregulated* gene sets only.
1. Save plots and tables to `results` directory.

## Paths and Packages

```{r paths_packages}
library(annotables)
library(limma)
library(pathwaze)
library(tidyverse)

# Set the default ggplot theme to "theme_bw" with specified base font size
theme_set(theme_bw(20))

# If on Domino save to results dir
RESULTS_DIR <- "."
if (Sys.getenv("DOMINO_WORKING_DIR") != "") RESULTS_DIR <- "../results"
```

## Data Sources

```{r load_data}
baselineFit <- readRDS("/stash/results/dev/rossmacp/P01376_CA209009_Response_Manuscript_Figures/checkmate9_CD3TCRsig_baseline_limma.rds")

day28Fit <- readRDS("/stash/results/dev/rossmacp/P01376_CA209009_Response_Manuscript_Figures/checkmate9_CD3TCRsig_Day28_Response_limma.rds")

probeset <- readRDS("/stash/results/dev/rossmacp/P01376_CA209009_Response_Manuscript_Figures/probeset_annotation.rds") %>%
  mutate(entrez_unique=str_extract(Probeset, "^\\d+")) %>%
  select(Probeset, entrez_unique)

# Load gene sets from MSigDB
h <- getMSigDBCollection("h")
c2 <- getMSigDBCollection("c2")
```

## Gene set enrichment

We perform gene set enrichment using the [`limma::camera` method](https://www.ncbi.nlm.nih.gov/pubmed/22638577) method with MSigDB C2 and Hallmark gene sets.

```{r gene_set_enrichment, warning=FALSE}
runGSE <- function(fit, filePrefix, collection, unadjustedPThreshold=1, netDown, w, h, ...) {
  # Helper function to execute gene set enrichment.
  # Params:
  #   fit = limma fit object
  #   filePrefix = output file prefix
  #   collection = GeneSetCollection object for DB of gene sets/pathways
  #   unadjustedPThreshold = When no genes are significant, we may want to only
  #                          consider significant genes *before* multiple testing
  #                          correction [default=1]
  #   netDown = boolean indicating if we are interested in gene sets with a net
  #             downregulation (ignoring enriched, but upregulated, gene sets)
  #   w = figure width
  #   h = figure height
  #   ... = arguments passed to pathwaze::runCameraPR()
  
  # Extract t-statistic and p-values and convert probeset IDs to Entrez IDs
  tt <- topTable(fit, number=Inf) %>%
    rownames_to_column("Probeset") %>%
    filter(P.Value <= unadjustedPThreshold)
  
  tDat <- tt %>%
    select(Probeset, t) %>%
    left_join(probeset, by="Probeset") %>%
    select(-Probeset) %>%
    column_to_rownames("entrez_unique") %>%
    as.matrix()
  
  pDat <- tt %>%
    select(Probeset, P.Value) %>%
    left_join(probeset, by="Probeset") %>%
    select(-Probeset) %>%
    column_to_rownames("entrez_unique") %>%
    as.matrix()
  
  # Compute enrichment
  gse <- runCameraPR("human", collection, tDat, pDat, ...)
  
  if (netDown) {
    # Only consider the gene sets that are net downregulated
  	# Add Symbol and description using annotables
    gse$enriched <- filter(gse$enriched, Direction == "Down")
    gse$setDetails <- semi_join(gse$setDetails, gse$enriched, by="set_name") %>%
    	mutate(entrez_9606 = as.numeric(entrez_9606)) %>%
    	left_join(select(grch37, entrez, symbol, description), 
    			  by=c("entrez_9606"="entrez"))
    
  }
  
  # The fraction in the plotting function below is the ADJ P value cutoff for plotting
  ggsave(
    filename=file.path(RESULTS_DIR, paste0(filePrefix, "_plot.png")),
    plot=geneSetEnrichmentPlot(gse$setDetails, 0.1),
    device="png", width=w, height=h)
  
  write.csv(gse$enriched,
            file=file.path(RESULTS_DIR, paste0(filePrefix, "_table.csv")),
            row.names=FALSE)
  
  write.csv(gse$setDetails,
            file=file.path(RESULTS_DIR, paste0(filePrefix, "_table_detail.csv")),
            row.names=FALSE)
}

# Only gene sets that are net downregulated for baseline
runGSE(baselineFit, "pathwaze_CD3TCRsig_baseline_Hallmark", h, netDown=TRUE, w=7, h=3)
runGSE(baselineFit, "pathwaze_CD3TCRsig_baseline_C2", c2, netDown=TRUE, w=7, h=12)
```

Nothing significantly DE in the Day28 contrast, so relax the FDR threshold.

Relaxed to FDR < 0.5 yields `r filter(topTable(day28Fit, number=Inf), adj.P.Val < 0.5) %>% nrow` DE genes.

Relaxed to FDR < 0.6 yields `r filter(topTable(day28Fit, number=Inf), adj.P.Val < 0.6) %>% nrow` DE genes.

We will continue with a threshold of 0.6 for the gene set enrichment.

```{r day28_gse, warning=FALSE}
runGSE(day28Fit, "pathwaze_CD3TCRsigHi_Day28_Hallmark", h, unadjustedPThreshold=0.05,
       netDown=FALSE, w=7, h=6, deSigLevel=0.99)
runGSE(day28Fit, "pathwaze_CD3TCRsigHi_Day28_C2", c2, unadjustedPThreshold=0.05,
       netDown=FALSE, w=7, h=6, deSigLevel=0.99)
```

## `R` session information

```{r, echo_session_info}
sessionInfo()
```
