---
title: "CA209009 baseline Geneset Scores and ccrcc group calculation"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Clear the environment
rm(list = ls())

# Free up memory by forcing garbage collection
invisible(gc())  

# Manually set the seed to an arbitrary number for consistency in reports
myseed <- 9
```

## Procedure

1. Create each of the geneset probeset lists
1. Get RMA expression values for baseline samples for each of the lists
1. Calculate score for each list
1. Merge score into sample annotation
1. Calculate ccrcc grouping using top ccrcc list score for patient
1. Calculate ccrcc grouping using clusters
1. Merge ccrcc grouping into sample annotation
1. Save annotation for analytical work

## Paths and Packages

```{r paths_packages}
# Provide paths
data_dir <- "./data/import"
results_dir <- "./results"
work_dir <- "./work"


## Load packages ##

#tidyverse  bundles: ggplot2, dplyr, tidyr, readr, purrr and tibble
suppressPackageStartupMessages(library(tidyverse))
library(ComplexHeatmap)
library(circlize)
library(limma)
library(knitr)
library(ggpubr)
library(broom)
```


```{r palettes}

ccrcc_colors = c("ccrcc1"= "black",
							  					  "ccrcc2"= "red",
							  					  "ccrcc3"= "grey",
							  					  "ccrcc4"= "goldenrod")


```

```{r functions}

source("./code/ggplot_theme_dj_prm.R")
#' Custom ggplot theme with bolded text for easier legibility


## Functions by Peter Szabo

contrastTable <- function(object, ...)
  UseMethod("contrastTable")

contrastTable.coxph <- function(object, linmat, level = 0.95, df = 1, rnames =
                                  dimnames(linmat)[[1]]) {
  #
  # DESCRIPTION:
  #    Computes point estimates and confidence intervals, and Wald
  #    test statistics and p-values, for specified contrasts (or
  #    more generally, linear combinations) of coefficients from a
  #    fitted Cox model (coxph object). Typically these contrasts
  #    will represent some sort of hazard ratio (HR).
  #
  #    This is a method for the generic function contrastTable for
  #    objects inheriting from class "coxph".
  #
  # ARGUMENTS:
  # object  a coxph object
  # linmat  numeric matrix with one row per contrast and one column
  #    corresponding to each coefficient in object.
  # level  desired confidence level. Default is 0.95, giving 95%
  #    confidence intervals for the contrasts.
  # df  a vector of positive integers, giving the degrees of
  #    freedom for the chi-squared distribution used for computing
  #    the Wald test p-values. The default is 1. Must have length
  #    one, in which case that value is used for all tests, or
  #    length nrow(linmat). If you're unsure what value to use for
  #    a particular contrast, consult a statistician. If you ARE a
  #    statistician, consult a different statistician. Has no
  #    influence on HR point and interval estimates.
  # rnames  character vector giving names of the requested
  #    contrasts, used as rownames in the return value. Default is
  #    row names of linmat.
  #
  # VALUE:
  #    a numeric matrix with one row per contrast, and columns
  #    "HR", "logHR", "SE.logHR", "lo.logHR", "up.logHR", "level",
  #    "WaldStat", and "Pvalue".
  #
  ###
  # Argument checking
  ###
  if (length(rnames) == 0) rnames <- character(0)
  if (!is.matrix(linmat))
    stop("linmat must be a matrix.")
  ncon <- nrow(linmat)
  if (ncon == 0) return(NULL)
  b <- coef(object)
  if (ncol(linmat) != length(b))
    stop(paste("linmat must have", length(b), "columns."))
  if (length(df) != 1 && length(df) != ncon)
    stop(paste("df must have length 1 or ", ncon, ".", sep = ""))
  if (length(level) != 1)
    stop("level must have length 1.")
  ###
  # Compute useful stuff
  ###   
  loghr <- linmat %*% b
  hrest <- exp(loghr)
  se.loghr <- sqrt(diag(linmat %*% object$var %*% t(linmat)))
  qqq <- qnorm( 1 - ( (1 - level)/2 ) )
  lo <- loghr - qqq * se.loghr
  hi <- loghr + qqq * se.loghr
  waldval <- loghr/se.loghr
  pwald <- 1 - pchisq(waldval^2, df)
  ###
  # Wrap it up pretty and go home
  ###
  cnames <- c( "HR", "logHR", "SE.logHR", "lo.logHR", "up.logHR", "level",
               "WaldStat", "Pvalue" )
  level <- rep(level, ncon)
  out <- c(hrest, loghr, se.loghr, lo, hi, level, waldval, pwald)
  out <- array(out, c(ncon, round(length(out)/ncon)), list(rnames, cnames))
  out
}





contrastTable.glm <- function(object, linmat, level = 0.95, df = 1, rnames =
                                dimnames(linmat)[[1]]) {
  #
  # DESCRIPTION:
  #    Computes point estimates and confidence intervals, and Wald
  #    test statistics and p-values, for specified contrasts (or
  #    more generally, linear combinations) of coefficients from a
  #    fitted Cox model (coxph object). Typically these contrasts
  #    will represent some sort of hazard ratio (HR).
  #
  #    This is a method for the generic function contrastTable for
  #    objects inheriting from class "coxph".
  #
  # ARGUMENTS:
  # object  a coxph object
  # linmat  numeric matrix with one row per contrast and one column
  #    corresponding to each coefficient in object.
  # level  desired confidence level. Default is 0.95, giving 95%
  #    confidence intervals for the contrasts.
  # df  a vector of positive integers, giving the degrees of
  #    freedom for the chi-squared distribution used for computing
  #    the Wald test p-values. The default is 1. Must have length
  #    one, in which case that value is used for all tests, or
  #    length nrow(linmat). If you're unsure what value to use for
  #    a particular contrast, consult a statistician. If you ARE a
  #    statistician, consult a different statistician. Has no
  #    influence on HR point and interval estimates.
  # rnames  character vector giving names of the requested
  #    contrasts, used as rownames in the return value. Default is
  #    row names of linmat.
  #
  # VALUE:
  #    a numeric matrix with one row per contrast, and columns
  #    "HR", "logHR", "SE.logHR", "lo.logHR", "up.logHR", "level",
  #    "WaldStat", and "Pvalue".
  #
  ###
  # Argument checking
  ###
  if (length(rnames) == 0) rnames <- character(0)
  if (!is.matrix(linmat))
    stop("linmat must be a matrix.")
  ncon <- nrow(linmat)
  if (ncon == 0) return(NULL)
  b <- coef(object)
  if (ncol(linmat) != length(b))
    stop(paste("linmat must have", length(b), "columns."))
  if (length(df) != 1 && length(df) != ncon)
    stop(paste("df must have length 1 or ", ncon, ".", sep = ""))
  if (length(level) != 1)
    stop("level must have length 1.")
  ###
  # Compute useful stuff
  ###   
  logor <- linmat %*% b
  hrest <- exp(logor)
  se.logor <- sqrt(diag(linmat %*% summary(object)$cov.scaled %*% t(linmat)))
  qqq <- qnorm( 1 - ( (1 - level)/2 ) )
  lo <- logor - qqq * se.logor
  hi <- logor + qqq * se.logor
  waldval <- logor/se.logor
  pwald <- 1 - pchisq(waldval^2, df)
  ###
  # Wrap it up pretty and go home
  ###
  cnames <- c( "OR", "logOR", "SE.logOR", "lo.logOR", "up.logOR", "level",
               "WaldStat", "Pvalue" )
  level <- rep(level, ncon)
  out <- c(hrest, logor, se.logor, lo, hi, level, waldval, pwald)
  out <- array(out, c(ncon, round(length(out)/ncon)), list(rnames, cnames))
  out
}


```

## Data Sources

```{r loaddata}

# Expression values
rma_file <- paste(data_dir, "CA209009-tumorAffy-HGU219_HS_ENTREZG.rma", sep = "/")
rma <- read_tsv(rma_file)

# Make rma Colnames match "Assay.Name" in sdrf annotation
colnames(rma) <- sub("ENTREZG-", "", colnames(rma))
rma <- rename(rma, Probeset = X1)

# SDRF (Sample and Data Relationship Format) file from Array Express
# NB move from stash when finalized.
sdrf_file <- paste(data_dir, "E-MTAB-3218 sdrf_response.txt", sep = "/")
sdrf <- read_tsv(sdrf_file)
sdrf$Response20pct <- as.factor(sdrf$Response20pct)
sdrf$individual <- as.factor(sdrf$individual)
sdrf$clinical.history <- as.factor(sdrf$clinical.history)

# Affymetrix probeset to Gene annotation
#JFC jimhester actually deliberately made it wipe the data from a column if the first 1000 rows are NA
#because it now uses 1000 rows to guess type = logical, when it isn't.
probeset_file <- paste(data_dir, "U219_BrainArray_Locus_Symbol_IRIS.txt", sep = "/")
probeset <- read_tsv(probeset_file,guess_max = 18571)
```

```{r geneset_vectors}
# Object containing Probesets for Geneset of T cell receptor transcripts
tcellprobesets <- probeset %>%
	filter(!is.na(probeset$Geneset_Tcell)) %>%
	pull(Probeset)

# # Object containing 88 Probesets for Geneset of ELVIDGE_HIF1A_TARGETS_DN
# hif1a_probesets <- probeset %>%
# 	filter(!is.na(probeset$Geneset_ELVIDGE_HIF1A_TARGETS_DN)) %>%
# 	pull(Probeset)

# McDermott IMmotion 150 genesets
#McDermott et al, Nature Medicine volume 24, pages 749–757 (2018)

Geneset_Angio_probesets <- probeset %>%
	filter(Geneset_IMmotion150 == "McDermott_Angio") %>%
	pull(Probeset) 


Geneset_Myeloid_probesets<- probeset %>%
	filter(Geneset_IMmotion150 == "McDermott_Myeloid") %>%
	pull(Probeset) 

Geneset_Teff_probesets<- probeset %>%
	filter(Geneset_IMmotion150 == "McDermott_Teff") %>%
	pull(Probeset) 

## Merck 18
Merckup<- probeset %>%
	filter(!is.na(probeset$Geneset_Merck18) & probeset$Geneset_Merck18=="Merck_Up") %>%
	pull(Probeset) 

Merckdown<- probeset %>%
	filter(!is.na(probeset$Geneset_Merck18) & probeset$Geneset_Merck18=="Merck_Down") %>%
	pull(Probeset) 

#Fuhrman
# Table S2 from Thibodeau 
#10.1016/j.urolonc.2015.11.001

Fuhup<- probeset %>%
	filter(!is.na(probeset$Geneset_Fuhrman) & probeset$Geneset_Fuhrman== "Up") %>%
	pull(Probeset) 

Fuhdown<- probeset %>%
	filter(!is.na(probeset$Geneset_Fuhrman) & probeset$Geneset_Fuhrman== "Down") %>%
	pull(Probeset) 

# Corvus adenosine signatures

adenosineNanostring <- probeset %>%
	filter(grepl("Nano",probeset$Geneset_Adenosine)) %>%
	pull(Probeset)  
	
adenosineRNAseq <- probeset %>%
	filter(grepl("RNA",probeset$Geneset_Adenosine)) %>%
	pull(Probeset) 

adenosineShort <- probeset %>%
	filter(grepl("Short",probeset$Geneset_Adenosine)) %>%
	pull(Probeset)  


#ccrcc1 score
ccrcc1up<- probeset %>%
	filter(!is.na(probeset$Geneset_ccRCC) & probeset$Geneset_ccRCC== "ccrcc1.up") %>%
	pull(Probeset) 

ccrcc1down<- probeset %>%
	filter(!is.na(probeset$Geneset_ccRCC) & probeset$Geneset_ccRCC== "ccrcc1.dn") %>%
	pull(Probeset) 


#ccrcc2 score
ccrcc2up<- probeset %>%
	filter(!is.na(probeset$Geneset_ccRCC) & probeset$Geneset_ccRCC== "ccrcc2.up") %>%
	pull(Probeset) 

ccrcc2down<- probeset %>%
	filter(!is.na(probeset$Geneset_ccRCC) & probeset$Geneset_ccRCC== "ccrcc2.dn") %>%
	pull(Probeset) 

#ccrcc3 score
ccrcc3up<- probeset %>%
	filter(!is.na(probeset$Geneset_ccRCC) & probeset$Geneset_ccRCC== "ccrcc3.up") %>%
	pull(Probeset) 

ccrcc3down<- probeset %>%
	filter(!is.na(probeset$Geneset_ccRCC) & probeset$Geneset_ccRCC== "ccrcc3.dn") %>%
	pull(Probeset) 


#ccrcc4 score
ccrcc4up<- probeset %>%
	filter(!is.na(probeset$Geneset_ccRCC) & probeset$Geneset_ccRCC== "ccrcc4.up") %>%
	pull(Probeset) 

ccrcc4down<- probeset %>%
	filter(!is.na(probeset$Geneset_ccRCC) & probeset$Geneset_ccRCC== "ccrcc4.dn") %>%
	pull(Probeset)

#BMS93
bmsup<- probeset %>%
	filter(!is.na(probeset$Geneset_CM009) & probeset$Geneset_CM009== "CM009_up") %>%
	pull(Probeset) 

bmsdown<- probeset %>%
	filter(!is.na(probeset$Geneset_CM009) & probeset$Geneset_CM009== "CM009_down") %>%
	pull(Probeset)

#EMT stroma
emt <- probeset %>%
	filter(!is.na(probeset$Geneset_EMTstroma)) %>%
	pull(Probeset)

#Javelin
javelin <- probeset %>%
	filter(!is.na(probeset$Geneset_Javelin)) %>%
	pull(Probeset)

```

RMA expression data is read from

+ *`r rma_file `*

Sample annotation is read from

+ *`r sdrf_file `*

Probesets annotated with Signature Genesets are read from:

+ *`r probeset_file `*

Geneset sources are documented in: 

+ Probeset_Geneset_Annotation.Rmd


## Filter Baseline Expression dataset

Filter the RMA data to the baseline samples, and then to the probesets for each geneset.
 
```{r baseline_sets}
# Baseline Subject sample Annotation
sdrfScreen <- sdrf %>%
	filter(biopsy.timepoint == "Screen", Response20pct != "NE")

# Select 56 Screen Array columns plus Probeset column
# 'Select' gets Assay.Name cols in the same order as the sdrf annotation
# Order is Vital for ComplexHeatmap!! Vital for limma!

rmascreen <- select(rma, one_of("Probeset",sdrfScreen$Assay.Name))


# Subset rmascreen data to each of the score genesets
rmaTcell <- rmascreen[rmascreen$Probeset %in% tcellprobesets,]
#rmahif1a <- rmascreen[rmascreen$Probeset %in% hif1a_probesets,]
rmaAngio <- rmascreen[rmascreen$Probeset %in% Geneset_Angio_probesets,]
rmaMyeloid <- rmascreen[rmascreen$Probeset %in% Geneset_Myeloid_probesets,]
rmaTeff <- rmascreen[rmascreen$Probeset %in% Geneset_Teff_probesets,]
rmaMerckup <- rmascreen[rmascreen$Probeset %in% Merckup,]
rmaMerckdown <- rmascreen[rmascreen$Probeset %in% Merckdown,]
rmaFuhup <- rmascreen[rmascreen$Probeset %in% Fuhup,]
rmaFuhdown <- rmascreen[rmascreen$Probeset %in% Fuhdown,]
rmaadenosineNano <- rmascreen[rmascreen$Probeset %in% adenosineNanostring,]
rmaadenosineRNA <- rmascreen[rmascreen$Probeset %in% adenosineRNAseq,]
rmaadenosineShort <- rmascreen[rmascreen$Probeset %in% adenosineShort,]
rmaccrcc1up <- rmascreen[rmascreen$Probeset %in% ccrcc1up,]
rmaccrcc1down <- rmascreen[rmascreen$Probeset %in% ccrcc1down,]
rmaccrcc2up <- rmascreen[rmascreen$Probeset %in% ccrcc2up,]
rmaccrcc2down <- rmascreen[rmascreen$Probeset %in% ccrcc2down,]
rmaccrcc3up <- rmascreen[rmascreen$Probeset %in% ccrcc3up,]
rmaccrcc3down <- rmascreen[rmascreen$Probeset %in% ccrcc3down,]
rmaccrcc4up <- rmascreen[rmascreen$Probeset %in% ccrcc4up,]
rmaccrcc4down <- rmascreen[rmascreen$Probeset %in% ccrcc4down,]
rmabmsup <- rmascreen[rmascreen$Probeset %in% bmsup,]
rmabmsdown <- rmascreen[rmascreen$Probeset %in% bmsdown,]
rmaemt <- rmascreen[rmascreen$Probeset %in% emt,]
rmajavelin <- rmascreen[rmascreen$Probeset %in% javelin,]

```
 
## Calculation of Signature Scores

RMA values for each probeset in a Geneset are Z-scored across the screen samples. For each Sample, the median of the Z-scores is the final score.  

When a score is composed of 'up' and 'down' components, the z-score for the 'down' genes is calculated seperately, then the 'down' Z-scores are multiplied by -1 to flip the sign. For negative 'down' scores this gives a positive value, and for positive 'down' scores this gives a negative value. Then the data tables are merged and the median score is taken. Method weights the contribution of each gene equally. The median -score could be from a positive or negative gene.

Scores are then merged into the Sample Annotation table for the Screen Samples.

```{r Calculate_score}

##### CD3score
# Transpose rma data to columns, Z score, transpose back to rows, get median, transpose to a sample level annotation
tcellscore<- as.data.frame(t(scale(t(rmaTcell[,-1])))) %>%
	summarize_all(funs(median)) %>%
	gather(key = "Assay.Name", value = "CD3TCR.Score")


# merge score into SDRF sample annotation
sdrfScreen <- right_join(sdrfScreen, tcellscore,
								  by = "Assay.Name")

###### angioScore
# Transpose rma data to columns, Z score, transpose back to rows, get median, transpose to a sample level annotation
angioScore<- as.data.frame(t(scale(t(rmaAngio[,-1])))) %>%
	summarize_all(funs(median)) %>%
	gather(key = "Assay.Name", value = "IM150_Angio.Score")


# merge score into SDRF sample annotation
sdrfScreen <- right_join(sdrfScreen, angioScore,
								  by = "Assay.Name")
##### myeloidScore
# Transpose rma data to columns, Z score, transpose back to rows, get median, transpose to a sample level annotation
myeloidScore<- as.data.frame(t(scale(t(rmaMyeloid[,-1])))) %>%
	summarize_all(funs(median)) %>%
	gather(key = "Assay.Name", value = "IM150_MyeloidInfl.Score")


# merge score into SDRF sample annotation
sdrfScreen <- right_join(sdrfScreen, myeloidScore,
								  by = "Assay.Name")


##### teffScore
# Transpose rma data to columns, Z score, transpose back to rows, get median, transpose to a sample level annotation
teffScore<- as.data.frame(t(scale(t(rmaTeff[,-1])))) %>%
	summarize_all(funs(median)) %>%
	gather(key = "Assay.Name", value = "IM150_Teff.Score")


# merge score into SDRF sample annotation
sdrfScreen <- right_join(sdrfScreen, teffScore,
								  by = "Assay.Name")

##### merck18Score
#Do scores by calculating up AND down, flip sign on down then merge, then taking median
merckUpScore <- as.data.frame(t(scale(t(rmaMerckup[,-1]))))
merckDownScore <- -1*(as.data.frame(t(scale(t(rmaMerckdown[,-1])))))

merckScore <- rbind(merckUpScore,merckDownScore) %>%
	summarize_all(funs(median)) %>%
	gather(key = "Assay.Name", value = "Merck18.Score")

# merge score into SDRF sample annotation
sdrfScreen <- right_join(sdrfScreen, merckScore,
								  by = "Assay.Name")




#####Fuhrman Score
fuhUpScore<- as.data.frame(t(scale(t(rmaFuhup[,-1]))))
fuhDownScore <- -1*(as.data.frame(t(scale(t(rmaFuhdown[,-1])))))

fuhScore <- rbind(fuhUpScore,fuhDownScore) %>%
	summarize_all(funs(median)) %>%
	gather(key = "Assay.Name", value = "Fuhrman.Score")

# merge score into SDRF sample annotation
sdrfScreen <- right_join(sdrfScreen, fuhScore,
								  by = "Assay.Name")



#####
#Corvus adenosine short
# Transpose rma data to columns, Z score, transpose back to rows, get median, transpose to a sample level annotation
adenosineShortScore <- as.data.frame(t(scale(t(rmaadenosineShort[,-1])))) %>%
	summarize_all(funs(median)) %>%
	gather(key = "Assay.Name", value = "Ad_Short.Score")

# merge score into SDRF sample annotation
sdrfScreen <- right_join(sdrfScreen, adenosineShortScore,
								  by = "Assay.Name")

#Corvus adenosine NANOSTRING
# Transpose rma data to columns, Z score, transpose back to rows, get median, transpose to a sample level annotation
adenosineNanoScore <- as.data.frame(t(scale(t(rmaadenosineNano[,-1])))) %>%
	summarize_all(funs(median)) %>%
	gather(key = "Assay.Name", value = "Adenosine.Score")

# merge score into SDRF sample annotation
sdrfScreen <- right_join(sdrfScreen, adenosineNanoScore,
								  by = "Assay.Name")

#Corvus adenosine RNAseq
# Transpose rma data to columns, Z score, transpose back to rows, get median, transpose to a sample level annotation
adenosineRNAseqScore <- as.data.frame(t(scale(t(rmaadenosineRNA[,-1])))) %>%
	summarize_all(funs(median)) %>%
	gather(key = "Assay.Name", value = "Ad_RNAseq.Score")

# merge score into SDRF sample annotation
sdrfScreen <- right_join(sdrfScreen, adenosineRNAseqScore,
								  by = "Assay.Name")

##### ccrcc1 Score
ccrcc1UpScore<- as.data.frame(t(scale(t(rmaccrcc1up[,-1]))))
ccrcc1DownScore <- -1*(as.data.frame(t(scale(t(rmaccrcc1down[,-1])))))

ccrcc1Score <- rbind(ccrcc1UpScore,ccrcc1DownScore) %>%
	summarize_all(funs(median)) %>%
	gather(key = "Assay.Name", value = "ccrcc1.Score")

# merge score into SDRF sample annotation
sdrfScreen <- right_join(sdrfScreen, ccrcc1Score,
								  by = "Assay.Name")
##### ccrcc2 Score
ccrcc2UpScore<- as.data.frame(t(scale(t(rmaccrcc2up[,-1]))))
ccrcc2DownScore <- -1*(as.data.frame(t(scale(t(rmaccrcc2down[,-1])))))

ccrcc2Score <- rbind(ccrcc2UpScore,ccrcc2DownScore) %>%
	summarize_all(funs(median)) %>%
	gather(key = "Assay.Name", value = "ccrcc2.Score")

# merge score into SDRF sample annotation
sdrfScreen <- right_join(sdrfScreen, ccrcc2Score,
								  by = "Assay.Name")

##### ccrcc3 Score
ccrcc3UpScore<- as.data.frame(t(scale(t(rmaccrcc3up[,-1]))))
ccrcc3DownScore <- -1*(as.data.frame(t(scale(t(rmaccrcc3down[,-1])))))

ccrcc3Score <- rbind(ccrcc3UpScore,ccrcc3DownScore) %>%
	summarize_all(funs(median)) %>%
	gather(key = "Assay.Name", value = "ccrcc3.Score")

# merge score into SDRF sample annotation
sdrfScreen <- right_join(sdrfScreen, ccrcc3Score,
								  by = "Assay.Name")

##### ccrcc4 Score
ccrcc4UpScore<- as.data.frame(t(scale(t(rmaccrcc4up[,-1]))))
ccrcc4DownScore <- -1*(as.data.frame(t(scale(t(rmaccrcc4down[,-1])))))

ccrcc4Score <- rbind(ccrcc4UpScore,ccrcc4DownScore) %>%
	summarize_all(funs(median)) %>%
	gather(key = "Assay.Name", value = "ccrcc4.Score")

# merge score into SDRF sample annotation
sdrfScreen <- right_join(sdrfScreen, ccrcc4Score,
								  by = "Assay.Name")


#####bms.Score
bmsUpScore<- as.data.frame(t(scale(t(rmabmsup[,-1]))))
bmsDownScore <- -1*(as.data.frame(t(scale(t(rmabmsdown[,-1])))))

bmsScore <- rbind(bmsUpScore,bmsDownScore) %>%
	summarize_all(funs(median)) %>%
	gather(key = "Assay.Name", value = "BMS.Score")

# merge score into SDRF sample annotation
sdrfScreen <- right_join(sdrfScreen, bmsScore,
								  by = "Assay.Name")

##### EMTstromaScore
# Transpose rma data to columns, Z score, transpose back to rows, get median, transpose to a sample level annotation
emtScore<- as.data.frame(t(scale(t(rmaemt[,-1])))) %>%
	summarize_all(funs(median)) %>%
	gather(key = "Assay.Name", value = "EMTstroma.Score")


# merge score into SDRF sample annotation
sdrfScreen <- right_join(sdrfScreen, emtScore,
								  by = "Assay.Name")

##### JavelinScore
# Transpose rma data to columns, Z score, transpose back to rows, get median, transpose to a sample level annotation
javelinScore<- as.data.frame(t(scale(t(rmajavelin[,-1])))) %>%
	summarize_all(funs(median)) %>%
	gather(key = "Assay.Name", value = "Javelin.Score")


# merge score into SDRF sample annotation
sdrfScreen <- right_join(sdrfScreen, javelinScore,
								  by = "Assay.Name")
```



## ccrcc group assignment

### ccrccTopScore: Use the max score values to assign ccrcc group

Here I am selecting "ccrccTopScore" as whichever score is the highest for that patient. 


```{r maxcol_ccrcc}

sdrfScreenccrcc <- sdrfScreen%>%
	select(ccrcc1.Score,
		   ccrcc2.Score,
		   ccrcc3.Score,
		   ccrcc4.Score)

ccrccTopScore<-colnames(sdrfScreenccrcc)[max.col(sdrfScreenccrcc,ties.method="first")]

sdrfScreen <- cbind(sdrfScreen,
				ccrccTopScore)

sdrfScreen$ccrccTopScore <- factor(sdrfScreen$ccrccTopScore,
						   levels = c("ccrcc1.Score",
		   "ccrcc2.Score",
		   "ccrcc3.Score",
		   "ccrcc4.Score"))

```

### Plot ccrccTopScore v. ccrcc4 

This plot confirms the Top Score was assigned correctly.

```{r boxplot_ccrccTopScore}


plotcount <- nrow(sdrfScreen)

boxplot <- sdrfScreen %>% 
			ggplot(aes(x = ccrccTopScore, y = ccrcc4.Score)) +
	geom_boxplot(outlier.shape = NA) +
	geom_point(aes(colour = Response20pct),
			   size = 3, 
			   position=position_jitterdodge(dodge.width=1, jitter.width = 0.2)) +
	scale_shape_manual(values=c(3,1)) +
	scale_colour_manual(name = 'Response',
						values = setNames(c('black','orange'),
										  c("Not20pct", "20pctDec")))+
	scale_y_continuous(breaks=seq(-1, 1, 1)) +
	labs(title = "CA209-009 Baseline Biopsy: ccrcc4 v. ccrccTopScore",
		 subtitle = paste("Subjects with Screen Affymetrix, N=",plotcount),
		 x = "ccrccTopScore Category",
		 y = "ccrcc4 signature") +
	theme(legend.position = "bottom",
		  axis.text=element_text(size=12),
		  axis.title=element_text(size=14,face="bold")) +
	guides(color=guide_legend(nrow=2,byrow=TRUE), 
		   shape = guide_legend(nrow=2,byrow=TRUE)) +
	stat_compare_means()

print(boxplot)
```

### ccrccCluster: hcluster to assign ccrcc

Cluster on the "70" gene expression data itself, not the 4 scores!! I tested several methods using CM214 data.

Ward.D2 performs the best at overlap, and is the method that proved most plausible in clustering ccrcc70 heatmaps. 

Using cutree(clusters, 4), the table has the 4 clusters in rows and the assignment by ccrccTopScore in the columns. 

After examining the table, I relabel the 4 clusters to reflect the ccrcc groups 1-4.

```{r cluster_ccrcc}


#filter the expression data to the 63 ccrcc70 probesets

ccrcc63<- probeset %>%
	filter(!is.na(probeset$Geneset_ccRCC)) %>%
	pull(Probeset) 


#Extract the expression data 
rmaccrcc <- rmascreen%>%
	filter(Probeset %in% ccrcc63)

# scaled_mat of expression values
# Z-score the data
scaled_mat = as.data.frame(scale(t(rmaccrcc[,-1])))
scaled_mat <- rownames_to_column(scaled_mat, var = "Assay.Name")

#merge to design to get in the same order

clusterdata <- sdrfScreen%>%
	select("Assay.Name")%>%
	left_join(scaled_mat,
			  by = "Assay.Name")


clusterswd2 <- hclust(dist(clusterdata[, 2:64]),
				   method = "ward.D2")

ccrccClusterwd2 <- cutree(clusterswd2, 4)


kable(table(ccrccClusterwd2, sdrfScreen$ccrccTopScore),
	  caption = "TopScore in columns, ward.D2 clusters in rows")

```

```{r add_ward2}

sdrfScreen <- cbind(sdrfScreen, ccrccClusterwd2)

sdrfScreen <- rename(sdrfScreen, ccrccCluster = ccrccClusterwd2)

sdrfScreen$ccrccCluster <- gsub("1","ccrccthree",sdrfScreen$ccrccCluster)
sdrfScreen$ccrccCluster <- gsub("2","ccrcc2",sdrfScreen$ccrccCluster)
sdrfScreen$ccrccCluster <- gsub("3","ccrcc1",sdrfScreen$ccrccCluster)
sdrfScreen$ccrccCluster <- gsub("4","ccrcc4",sdrfScreen$ccrccCluster)
sdrfScreen$ccrccCluster <- gsub("ccrccthree","ccrcc3",sdrfScreen$ccrccCluster)

```



## Outputs

```{r outputs}

#sdrf Dataset with signature scores
scores_file  <- paste0(results_dir, "/GEP_Table_BiopsyScreen_GEPSignatures_sdrf.txt")

write_tsv(sdrfScreen, scores_file)

```

Save the Screen sdrf with scores and ccrcc clusters to:

+*`r scores_file`*

