---
title: "Comparing ccrcc groups with BMS Baseline response predictors"
output: github_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Clear the environment
rm(list = ls())

# Free up memory by forcing garbage collection
invisible(gc())  

# Manually set the seed to an arbitrary number for consistency in reports
myseed <- 9

##_ Set knitr root directory to correspond to project working directory 
##_  setting based on structure with code in <project dir>/code
knitr::opts_knit$set(root.dir = here::here())
```

## Procedure

1. Load E-MTAB-3267 Probeset file containing BMS311 and ccrcc annotation
1. Load E-MTAB-3267 RMA dataset for 359 prbs
1. Load Brain array Probeset file containing BMS311 and ccrcc annotation
1. Load CA209009 screen RMA dataset
1. Heatmap of E-MTAB-3267 data for BMS predictor genes 
1. Heatmap of CA209009 data for ccrcc predictor genes


## Paths and Packages

```{r paths_packages}
# Provide paths
data_dir <- "./data/import"
results_dir <- "./results"
work_dir <- "./work"

## Load packages ##

#tidyverse  bundles: ggplot2, dplyr, tidyr, readr, purrr and tibble
suppressPackageStartupMessages(library(tidyverse))
library(knitr)

library(ComplexHeatmap)
library(circlize)

#packages for importing and processing MTAB3267 dataset
#library(ArrayExpress)
#library(affy)
#library(annotate)
#library(hugene10sttranscriptcluster.db)

```

## Data Sources

```{r processINSERM}

# #uses library("ArrayExpress")
# # should this be E-MTAB-3267
# AEset = ArrayExpress("E-MEXP-3267")
# 
# #uses library("affy")
# #Creates RMA processed data ExpressionSet (assayData,phenoData,featureData, protocolData,experimentData).
# AEsetnorm = rma(AEset)
# 
# #RMA values to a table, with the HuGene-1_0-st-v1 probeset IDs
# RMAtable <- exprs(AEsetnorm)
# 
# # Sample annotation 
# # Get the project's sample factors from ArrayExpress and add ccRCC class provided by Sylvie Job.
# 
# # Factors of interest to a table
# factors <- pData(AEsetnorm)[,1:15]
# 
# # make the rownames (i.e the cel file names) into their own named column
# factors <- factors %>% add_rownames()
# names(factors)[1] <- "CELfile"
# 
# # Additional factors were obtained from Sylvie Job
# ccrcc <- read.xlsx("./data/import/file_annot_EMTAB3267_for_Petra.xlsx", sheetIndex = 1)
# 
# # merge the additional factors from Sylvie into factors, using "Sample 1" etc
# allfactors <- left_join(factors, ccrcc, by = "Source.Name")
# 
# # Make Response into a factor with levels in the desired order
# allfactors$Characteristics.sunitinib.response. <- factor(allfactors$Characteristics.sunitinib.response., 
# levels  = c("PD", "CLINICAL BENEFIT", "SD", "PR"))
# 
# ## Probeset annotation
# # Affymetrix HuGene ST1.0 probesets annotation is obtained from Bioconductor library
# #uses library(annotate)
# #uses library(hugene10sttranscriptcluster.db)
# annodb <- "hugene10sttranscriptcluster.db"
# ID     <- featureNames(AEsetnorm)
# Symbol <- as.character(lookUp(ID, annodb, "SYMBOL"))
# Name   <- as.character(lookUp(ID, annodb, "GENENAME"))
# Entrez <- as.character(lookUp(ID, annodb, "ENTREZID"))
# HuGeneAnnot <- cbind.data.frame(ID,Symbol,Name,Entrez)
# 
# # Write these files to work directory and comment out the chunk
# 
# # write RMA data, Probeset annotation, sample factors to 3 files
# write.table(RMAtable, "./work/E_MEXP_3267.txt", sep="\t")
# write.table(HuGeneAnnot, "./work/HuGene1Anot.txt", sep="\t")
# write.table(allfactors, "./work/E_MEXP_3267_factors.txt", sep="\t")

```

EMTAB3267 is provided as a custom subset:

+ 51 samples of the 59 in ArrayExpress
+ 359 probesets, RMA data

```{r loaddata_EMTAB3267}

# Hugene Probeset file containing 359prbs with ccrcc and BMS annotation
hugene_file <- paste(data_dir, "HuGene1Anot_Predictors.txt", sep = "/")
hugeneprobeset <- read_tsv(hugene_file)

# RMA dataset for 359 prbs and 51 tumor with ccrcc in E-MTAB-3267
rma3267_file <- paste(data_dir, "E_MTAB_3267_51samples_359prbs_RMA.txt", sep = "/")
rma3267 <- read_tsv(rma3267_file)

# Sample annotation for 51 tumor with ccrcc in E-MTAB-3267
annot3267_file <- paste(data_dir, "E_MTAB_3267_51samples.txt", sep = "/")
annot3267 <- read_tsv(annot3267_file)

# RMA column order is same as sample row order

```

BMS screen biopsy data for the ccrcc70 genes is filtered from the files used for all analyses.

```{r load_data_BMS}

# Expression values
rma_file <- paste(data_dir, "CA209009-tumorAffy-HGU219_HS_ENTREZG.rma", sep = "/")
rma <- read_tsv(rma_file)

# Make rma Colnames match "Assay.Name" in sdrf annotation
colnames(rma) <- sub("ENTREZG-", "", colnames(rma))
rma <- dplyr::rename(rma, Probeset = X1)

# SDRF (Sample and Data Relationship Format) file from Array Express
# NB move from stash when finalized (eg PBRM1 added)
sdrf_file <- paste(data_dir, "E-MTAB-3218 sdrf_response.txt", sep = "/")
sdrf <- read_tsv(sdrf_file)
sdrf$Response20pct <- as.factor(sdrf$Response20pct)
sdrf$individual <- as.factor(sdrf$individual)
sdrf$clinical.history <- as.factor(sdrf$clinical.history)

# Affymetrix probeset to Gene annotation
probeset_file <- paste(data_dir, "U219_BrainArray_Locus_Symbol_IRIS.txt", sep = "/")
probeset <- read_tsv(probeset_file)

# Object containing ccrcc probesets
ccrcc_probesets <- probeset %>%
	filter(!is.na(probeset$Geneset_ccRCC)) %>%
	pull(Probeset)

```



## E-MTAB-3267 data analysis

We will use data for 51 samples of the 59 in ArrayExpress

+ 5 normal (non-tumor) removed
+ 3 samples without ccrcc annotation removed


## E-MTAB-3267 data for 311 BMS predictor genes

The 311 BMS predictor genes could be successfully mapped to 297 genes on the Affymetrix HuGene ST1.0 array. For simplicity these are referred to as "BMS311".

```{r heatmap_3267_BMS311}

# filter 297 probesets for just BMS311
hugene311 <- hugeneprobeset %>%
	filter(!is.na(BMS311)) %>%
	pull(ID)

# filter annotationfor just BMS311
annot311 <- hugeneprobeset %>%
	filter(!is.na(BMS311))

# filter the RMA data for the BMS 311 genes
rma311 <- rma3267 %>%
	filter(ID %in% hugene311)


# scaled_mat will be the heatmap of expression values
# Z-score the data
scaled_mat = t(scale(t(rma311[,-1])))
rownames(scaled_mat) = rma311$ID


# Provide column order as ccrcc4
# This works because RMA data columns in same order as sample annot rows
ordering = order(annot3267$ccrcc.subtype)

# Top heatmap annotation is sunitinib response and ccrcc
ha_top = HeatmapAnnotation(sunitinib=
						   	annot3267$`Characteristics.sunitinib.response.`,
						   ccrcc = annot3267$ccrcc.subtype,
						   col = list(sunitinib = 
						   		   	c("CLINICAL BENEFIT" = "goldenrod1",
						   		   	  "PR" = "goldenrod2",
						   		   	  "SD" = "darkolivegreen3",
						   		   	  "PD" = "black"),
						   		  ccrcc = 
						   		  	c(ccrcc1 = "darkolivegreen1",
						   		  	  ccrcc2 = "darkolivegreen2",
						   		  	  ccrcc3 = "darkolivegreen3",
						   		  	  ccrcc4 = "goldenrod2")),
						   annotation_height = unit(c(0.5, 0.5), "cm"))

#draw(ha_top,1:51)

heatmap_object = Heatmap(scaled_mat,
						 col = colorRamp2(c(-1, 0, 1), c("dodgerblue", "white", "firebrick")),
						 cluster_rows = TRUE,
						 km = 2,
						 show_row_names = FALSE,
						 cluster_columns = FALSE,
						 column_order = ordering,
						 column_title = "MTAB3267: 51 samples, 311 BMS Response, ccrcc sort",
						 show_column_names = FALSE,
						 width = unit(9, "cm"),
						 top_annotation = ha_top,
						 top_annotation_height = unit(2, "cm") 
						 ) +
	rowAnnotation(bms = annot311$BMS311,
				  col = list(bms = colorRamp2(c(0.66, 1, 1.5), 
				  							c("dodgerblue", "white", "firebrick"))),
				  width = unit(0.3, "cm"),
				  show_annotation_name = TRUE)


# Produce the heatmap and save to a file
heatmap_file <- paste0(results_dir,"/","GEP_Heatmap_EMTAB3267_BMS311.pdf")
pdf(file=heatmap_file)

draw(heatmap_object, padding = unit(c(4, 10, 4, 4), "mm"))

# Decorate the heatmap object, adding layers of texts or lines
decorate_annotation("sunitinib", 
					{grid.text("Sunitinib response",
							   unit(-2, "mm"), just = "right",
							   gp = gpar(fontsize = 8))})

decorate_annotation("ccrcc", 
					{grid.text("ccrcc class",
							   unit(-2, "mm"), just = "right",
							   gp = gpar(fontsize = 8))})

dev.off()
```

## E-MTAB-3267 data for 93 Top BMS predictor genes

The 311 BMS predictor genes (297 genes on the Affymetrix HuGene ST1.0 array) were filtered to 93 genes with >1.5-fold difference between response groups in CA209-009. For simplicity these are referred to as "BMS93".


```{r heatmap_3267_BMS93}

# filter 359 rows to 93 probesets for just BMS311 fold>1.5
hugene93 <- hugeneprobeset %>%
	filter(BMS311 >1.5 | BMS311 < 0.66) %>%
	pull(ID)

annot93 <- hugeneprobeset %>%
	filter(BMS311 >1.5 | BMS311 < 0.66)

# filter the RMA data for the BMS93 genes
rma93 <- rma3267 %>%
	filter(ID %in% hugene93)


# scaled_mat will be the heatmap of expression values
# Z-score the data
scaled_mat = t(scale(t(rma93[,-1])))
rownames(scaled_mat) = rma93$ID


# Provide column order as ccrcc4
# This works because RMA data columns in same order as sample annot rows
ordering = order(annot3267$ccrcc.subtype)

# Top heatmap annotation is sunitinib response and ccrcc
ha_top = HeatmapAnnotation(sunitinib=
						   	annot3267$`Characteristics.sunitinib.response.`,
						   ccrcc = annot3267$ccrcc.subtype,
						   col = list(sunitinib = 
						   		   	c("CLINICAL BENEFIT" = "goldenrod1",
						   		   	  "PR" = "goldenrod2",
						   		   	  "SD" = "darkolivegreen3",
						   		   	  "PD" = "black"),
						   		  ccrcc = 
						   		  	c(ccrcc1 = "darkolivegreen1",
						   		  	  ccrcc2 = "darkolivegreen2",
						   		  	  ccrcc3 = "darkolivegreen3",
						   		  	  ccrcc4 = "goldenrod2")),
						   annotation_height = unit(c(0.5, 0.5), "cm"))
#draw(ha_top,1:51)

heatmap_object = Heatmap(scaled_mat,
						 col = colorRamp2(c(-2, 0, 2), c("dodgerblue", "white", "firebrick")),
						 cluster_rows = TRUE,
						 km = 2,
						 show_row_names = FALSE,
						 cluster_columns = TRUE,
						 clustering_method_columns = "ward",
						 #column_order = ordering,
						 column_title = "MTAB3267: 51 samples for 93 BMS Response",
						 show_column_names = FALSE,
						 width = unit(9, "cm"),
						 top_annotation = ha_top,
						 top_annotation_height = unit(2, "cm") 
						 ) +
	rowAnnotation(bms = annot93$BMS311,
				  col = list(bms = colorRamp2(c(0.66, 1, 1.5), 
				  							c("dodgerblue", "white", "firebrick"))),
				  width = unit(0.3, "cm"),
				  show_annotation_name = TRUE)


# Produce the heatmap and save to a file
heatmap_file <- paste0(results_dir,"/","GEP_Heatmap_EMTAB3267_BMS93.pdf")
pdf(file=heatmap_file)

draw(heatmap_object, padding = unit(c(4, 10, 4, 4), "mm"))

# Decorate the heatmap object, adding layers of texts or lines
decorate_annotation("sunitinib", 
					{grid.text("Sunitinib response",
							   unit(-2, "mm"), just = "right",
							   gp = gpar(fontsize = 8))})

decorate_annotation("ccrcc", 
					{grid.text("ccrcc class",
							   unit(-2, "mm"), just = "right",
							   gp = gpar(fontsize = 8))})

dev.off()
```

## E-MTAB-3267 data for ccrcc predictor genes

The 70 ccrcc predictor genes could be successfully mapped to 63 probesets on the Hugene annotation. For simplicity these are referred to as "ccrcc70".

Since the 63 are a subset of the original predictive geneset, I double check they are able to group the E_MTAB_3267 dataset by ccRCC class.

```{r heatmap_3267_ccrcc_byclass}

# filter 359 rows to 63 probesets for ccrcc70
hugene70 <- hugeneprobeset %>%
	filter(!is.na(ccrcc)) %>%
	pull(ID)

#make the annotation and sort by Rcc group/type
annot70 <- hugeneprobeset %>%
	filter(!is.na(ccrcc)) %>%
 separate(ccrcc, into = c("ccRCCsubtype", "UpDown")) %>%
	arrange(ccRCCsubtype,desc(UpDown))


# filter the RMA data for the ccrcc70 genes
rma70 <- rma3267 %>%
	filter(ID %in% hugene70) 

#make the RMA data in order	by ccrcc group/type
order70 <- as.character(annot70$ID)
rma70 <- rma70[match(order70, as.character(rma70$ID)),]


# scaled_mat will be the heatmap of expression values
# Z-score the data
scaled_mat = t(scale(t(rma70[,-1])))
rownames(scaled_mat) = rma70$ID


# Provide column order as ccrcc
# This works because RMA data columns in same order as sample annot rows
ordering = order(annot3267$ccrcc.subtype)

# Top heatmap annotation is sunitinib response and ccrcc
ha_top = HeatmapAnnotation(sunitinib=
						   	annot3267$`Characteristics.sunitinib.response.`,
						   ccrcc = annot3267$ccrcc.subtype,
						   col = list(sunitinib = 
						   		   	c("CLINICAL BENEFIT" = "goldenrod1",
						   		   	  "PR" = "goldenrod2",
						   		   	  "SD" = "darkolivegreen3",
						   		   	  "PD" = "darkolivegreen4"),
						   		  ccrcc = 
						   		  	c(ccrcc1 = "darkolivegreen1",
						   		  	  ccrcc2 = "blue",
						   		  	  ccrcc3 = "darkolivegreen3",
						   		  	  ccrcc4 = "goldenrod2")),
						   annotation_height = unit(c(0.5, 0.5), "cm"))
#draw(ha_top,1:51)

heatmap_object = Heatmap(scaled_mat,
						 col = colorRamp2(c(-2, 0, 2), c("cyan", "white", "red")),
						 cluster_rows = FALSE,
						 show_row_names = FALSE,
						 split = annot70$ccRCCsubtype,
						 cluster_columns = FALSE,
						 column_order = ordering,
						 column_title = "MTAB3267: 51 samples for 70 ccrcc",
						 show_column_names = FALSE,
						 width = unit(8, "cm"),
						 top_annotation = ha_top,
						 top_annotation_height = unit(2, "cm") 
						 ) +
	rowAnnotation(class = annot70$ccRCCsubtype,
				  type = annot70$UpDown,
				  col = list(class = 
				  		   	c(ccrcc1 = "darkolivegreen1",
						   		  	  ccrcc2 = "blue",
						   		  	  ccrcc3 = "darkolivegreen3",
						   		  	  ccrcc4 = "goldenrod2"),
				  		   type = c(up  = "red",
				  		   		 dn = "blue")),
				  width = unit(0.5, "cm"),
				  show_annotation_name = TRUE)


# Produce the heatmap and save to a file
heatmap_file <- paste0(results_dir,"/","GEP_Heatmap_EMTAB3267_ccrcc70.pdf")
pdf(file=heatmap_file)

draw(heatmap_object, padding = unit(c(4, 10, 4, 4), "mm"))

# Decorate the heatmap object, adding layers of texts or lines
decorate_annotation("sunitinib", 
					{grid.text("Sunitinib response",
							   unit(-2, "mm"), just = "right",
							   gp = gpar(fontsize = 8))})

decorate_annotation("ccrcc", 
					{grid.text("ccrcc class",
							   unit(-2, "mm"), just = "right",
							   gp = gpar(fontsize = 8))})

dev.off()

```

```{r heatmap_3267_ccrcc_clustered}

# filter 359 rows to 63 probesets for ccrcc70
hugene70 <- hugeneprobeset %>%
	filter(!is.na(ccrcc)) %>%
	pull(ID)

#make the annotation and sort by Rcc group/type
annot70 <- hugeneprobeset %>%
	filter(!is.na(ccrcc)) %>%
 separate(ccrcc, into = c("ccRCCsubtype", "UpDown")) %>%
	arrange(ccRCCsubtype,desc(UpDown))


# filter the RMA data for the ccrcc70 genes
rma70 <- rma3267 %>%
	filter(ID %in% hugene70) 

#make the RMA data in order	by Rcc group/type
order70 <- as.character(annot70$ID)
rma70 <- rma70[match(order70, as.character(rma70$ID)),]


# scaled_mat will be the heatmap of expression values
# Z-score the data
scaled_mat = t(scale(t(rma70[,-1])))
rownames(scaled_mat) = rma70$ID


# Provide column order as ccrcc
# This works because RMA data columns in same order as sample annot rows
ordering = order(annot3267$ccrcc.subtype)

# Top heatmap annotation is sunitinib response and ccrcc
ha_top = HeatmapAnnotation(sunitinib=
						   	annot3267$`Characteristics.sunitinib.response.`,
						   ccrcc = annot3267$ccrcc.subtype,
						   col = list(sunitinib = 
						   		   	c("CLINICAL BENEFIT" = "goldenrod1",
						   		   	  "PR" = "goldenrod2",
						   		   	  "SD" = "darkolivegreen3",
						   		   	  "PD" = "darkolivegreen4"),
						   		  ccrcc = 
						   		  	c(ccrcc1 = "darkolivegreen1",
						   		  	  ccrcc2 = "blue",
						   		  	  ccrcc3 = "darkolivegreen3",
						   		  	  ccrcc4 = "goldenrod2")),
						   annotation_height = unit(c(0.5, 0.5), "cm"))
#draw(ha_top,1:51)

heatmap_object = Heatmap(scaled_mat,
						 col = colorRamp2(c(-2, 0, 2), c("cyan", "white", "red")),
						 cluster_rows = FALSE,
						 show_row_names = FALSE,
						 split = annot70$ccRCCsubtype,
						 cluster_columns = TRUE,
						 #column_order = ordering,
						 column_title = "MTAB3267: 51 samples for 70 ccrcc, Clustered",
						 show_column_names = FALSE,
						 width = unit(8, "cm"),
						 top_annotation = ha_top,
						 top_annotation_height = unit(2, "cm") 
						 ) +
	rowAnnotation(class = annot70$ccRCCsubtype,
				  type = annot70$UpDown,
				  col = list(class = 
				  		   	c(ccrcc1 = "darkolivegreen1",
						   		  	  ccrcc2 = "blue",
						   		  	  ccrcc3 = "darkolivegreen3",
						   		  	  ccrcc4 = "goldenrod2"),
				  		   type = c(up  = "red",
				  		   		 dn = "blue")),
				  width = unit(0.5, "cm"),
				  show_annotation_name = TRUE)


# Produce the heatmap and save to a file
heatmap_file <- paste0(results_dir,"/","GEP_Heatmap_EMTAB3267_ccrcc70_clustered.pdf")
pdf(file=heatmap_file)

draw(heatmap_object, padding = unit(c(4, 10, 4, 4), "mm"))

# Decorate the heatmap object, adding layers of texts or lines
decorate_annotation("sunitinib", 
					{grid.text("Sunitinib response",
							   unit(-2, "mm"), just = "right",
							   gp = gpar(fontsize = 8))})

decorate_annotation("ccrcc", 
					{grid.text("ccrcc class",
							   unit(-2, "mm"), just = "right",
							   gp = gpar(fontsize = 8))})

dev.off()

```

## CA209009 data for ccrcc predictor genes

The 70 ccrcc predictor genes could be successfully mapped to 65 genes on the BrainArray. However I am only using the 63 genes that were checked back on EMTAB3267. For simplicity these are referred to as "ccrcc70".


```{r ccrcc_baseline}
# Baseline Subject sample Annotation
sdrfScreen <- sdrf %>%
	filter(biopsy.timepoint == "Screen", Response20pct != "NE")

# Select 56 Screen Array columns plus Probeset column
# 'Select' gets Assay.Name cols in the same order as the sdrf annotation
# Order is Vital for ComplexHeatmap!! Vital for limma!

rmascreen <- dplyr::select(rma, one_of(c("Probeset",sdrfScreen$Assay.Name)))

# Subset rmascreen data to just the 63 ccrcc probesets 
rma70bms <- rmascreen[rmascreen$Probeset %in% ccrcc_probesets,]

```

The 63-genes in the ccrcc70 geneset successfully cluster the BMS data. 8 of the 13 nivolumab responders are in a 17 member group that looks like ccrcc4.

```{r heatmap_BMS_ccrcc70}


#make the bms annotation and sort by Rcc group/type
annot70bms <- probeset %>%
	filter(!is.na(probeset$Geneset_ccRCC)) %>%
 separate(Geneset_ccRCC, into = c("ccRCCsubtype", "UpDown")) %>%
	arrange(ccRCCsubtype,desc(UpDown))

#make the RMA data in order	by Rcc group/type
order70bms <- annot70bms$Probeset
rma70bms <- rma70bms[match(order70bms, rma70bms$Probeset),]


# scaled_mat will be the heatmap of expression values
# Z-score the data
scaled_mat = t(scale(t(rma70bms[,-1])))
rownames(scaled_mat) = rma70bms$Probeset


# Provide column order ?
# This works because RMA data columns in same order as sample annot rows
# ordering = order()

# Top heatmap annotation is nivo response and prior vegf
ha_top = HeatmapAnnotation(response = sdrfScreen$Response20pct,
							  #vegf = sdrfScreen$clinical.history,
						   	pbrm1 = sdrfScreen$PBRM1,
							  col = list(
							  	#vegf = c("PriorTherapy_Yes"="goldenrod2","PriorTherapy_No"="darkolivegreen4"),
							  		   response = c("Not20pct" ="black", "20pctDec"= "goldenrod2", "NE" = "white"),
							  pbrm1 = c("Missense_Mutation" = "black",
								  		  "Nonsense_Mutation" = "goldenrod2",
                                            "Frame_Shift_Del" = "goldenrod2" ,
								  		  "Frame_Shift_Ins" = "goldenrod2",
								  		  "Splice_Site" = "goldenrod2",
								  		  "WT" = "black"
								  		  )		   ),
							  na_col = "white",
							  annotation_height = unit(c(.5, .5), "cm"))
#draw(ha_top,1:51)

ha_bottom = HeatmapAnnotation(colname = anno_text(sdrfScreen$individual, rot = 90, just = "right", offset = unit(1, "npc") - unit(2, "mm"), 
							  					gp = gpar(fontsize = 7)),
							  annotation_height = unit(c(1), "cm"))




heatmap_object = Heatmap(scaled_mat,
						 col = colorRamp2(c(-2, 0, 2), c("dodgerblue", "white", "firebrick")),
						 cluster_rows = FALSE,
						 show_row_names = FALSE,
						 split = annot70bms$ccRCCsubtype,
						 cluster_columns = TRUE,
						 clustering_method_columns = "ward.D2",						 
						 #column_order = ordering,
						 column_title = 
						 	"CA209009: 56 screen samples for 70ccrcc, Clustered",
						 show_column_names = FALSE,
						 width = unit(8, "cm"),
						 top_annotation = ha_top,
						 top_annotation_height = unit(2, "cm"), 
						 bottom_annotation = ha_bottom,
						 bottom_annotation_height = unit(3, "cm")) +
	rowAnnotation(class = annot70bms$ccRCCsubtype,
				  type = annot70bms$UpDown,
				  col = list(class = 
				  		   	c(ccrcc1 = "darkolivegreen1",
						   		  	  ccrcc2 = "darkolivegreen2",
						   		  	  ccrcc3 = "darkolivegreen3",
						   		  	  ccrcc4 = "goldenrod2"),
				  		   type = c(up  = "red",
				  		   		 dn = "blue")),
				  width = unit(0.5, "cm"),
				  show_annotation_name = TRUE)


# Produce the heatmap and save to a file
heatmap_file <- paste0(results_dir,"/","GEP_Heatmap_BiopsyScreen_ccrcc70_clustered.pdf")
pdf(file=heatmap_file)

draw(heatmap_object, padding = unit(c(4, 10, 4, 4), "mm"))

# Decorate the heatmap object, adding layers of texts or lines
# decorate_annotation("vegf", 
# 					{grid.text("Prior VEGFi",
# 							   unit(-2, "mm"), just = "right",
# 							   gp = gpar(fontsize = 10))})
decorate_annotation("response", 
					{grid.text("Response",
							   unit(-2, "mm"), just = "right",
							   gp = gpar(fontsize = 10))})

decorate_annotation("pbrm1", 
					{grid.text("PBRM1 Status", unit(-2, "mm"), just = "right",
							   gp = gpar(fontsize = 8))})

dev.off()

```

The 21-gene ccrcc4 subset of the ccrcc70 geneset successfully clusters the BMS data. 8 of 13 nivolumab responders are in a 16 member group that looks like ccrcc4.

```{r heatmap_BMS_ccrcc21}


# subset the bms annotation to 21 genes in ccrcc4
annot21bms <- annot70bms %>%
	filter(ccRCCsubtype == "ccrcc4") %>%
	arrange(UpDown)

# subset the RMA data to ccrcc4
# Put cold genes on top to match other cluster

rma21bms <- rma70bms[rma70bms$Probeset %in% annot21bms$Probeset,]
rma21bms <- rma21bms[match(annot21bms$Probeset, rma21bms$Probeset),]


# scaled_mat will be the heatmap of expression values
# Z-score the data
scaled_mat = t(scale(t(rma21bms[,-1])))
rownames(scaled_mat) = rma21bms$Probeset


# Provide column order ?
# This works because RMA data columns in same order as sample annot rows
# ordering = order()

# Top heatmap annotation is nivo response and prior vegf
ha_top = HeatmapAnnotation(response = sdrfScreen$Response20pct,
							  #vegf = sdrfScreen$clinical.history,
							  col = list(#vegf = c("PriorTherapy_Yes"="goldenrod2","PriorTherapy_No"="darkolivegreen4"),
							  		   response = c("Not20pct" ="black", "20pctDec"= "goldenrod2", "NE" = "white")),
							  na_col = "white",
							  annotation_height = unit(c(.5), "cm"))
#draw(ha_top,1:51)

ha_bottom = HeatmapAnnotation(colname = anno_text(sdrfScreen$individual, rot = 90, just = "right", offset = unit(1, "npc") - unit(2, "mm"), 
							  					gp = gpar(fontsize = 7)),
							  annotation_height = unit(c(1), "cm"))




heatmap_object = Heatmap(scaled_mat,
						 col = colorRamp2(c(-2, 0, 2), c("dodgerblue", "white", "firebrick")),
						 cluster_rows = FALSE,
						 show_row_names = FALSE,
						 split = annot21bms$ccRCCsubtype,
						 cluster_columns = TRUE,
						 clustering_method_columns = "ward.D2",	
						 #column_order = ordering,
						 column_title = 
						 	"CA209009: 56 screen samples for 21ccrcc4, Clustered",
						 show_column_names = FALSE,
						 width = unit(10, "cm"),
						 top_annotation = ha_top,
						 top_annotation_height = unit(1, "cm"), 
						 bottom_annotation = ha_bottom,
						 bottom_annotation_height = unit(1, "cm")) +
	rowAnnotation(class = annot21bms$ccRCCsubtype,
				  type = annot21bms$UpDown,
				  col = list(class = 
				  		   	c(ccrcc1 = "darkolivegreen1",
						   		  	  ccrcc2 = "darkolivegreen2",
						   		  	  ccrcc3 = "darkolivegreen3",
						   		  	  ccrcc4 = "goldenrod2"),
				  		   type = c(up  = "red",
				  		   		 dn = "blue")),
				  width = unit(0.5, "cm"),
				  show_annotation_name = TRUE)


# Produce the heatmap and save to a file
heatmap_file <- paste0(results_dir,"/","GEP_Heatmap_BiopsyScreen_ccrcc4_clustered.pdf")
pdf(file=heatmap_file)

draw(heatmap_object, padding = unit(c(4, 10, 4, 4), "mm"))

# Decorate the heatmap object, adding layers of texts or lines
# decorate_annotation("vegf", 
# 					{grid.text("Prior VEGFi",
# 							   unit(-2, "mm"), just = "right",
# 							   gp = gpar(fontsize = 10))})

decorate_annotation("response", 
					{grid.text("Response",
							   unit(-2, "mm"), just = "right",
							   gp = gpar(fontsize = 10))})

dev.off()

```
